---
title: "Data statistics"
author: "Xinxiang Gao"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---
```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(widgetframe)
library(tidytext)

# INITALIZE CODE CHUNK OPTIONS
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{r}
source("data_process.R")
```

Here are some statistics for the final dataset.

## Sample Distribution {.tabset}

```{r}
plt <- function(col, name){
    ggplotly(ggplot(all_p_mr, aes_string(x = col, fill = col)) +
      geom_bar() +
      labs(title = paste("Bar plot of", name)) +
      coord_flip()) %>% 
    layout(legend = list(orientation = "h", y = -0.2, x = -0.3))
  }
```

### Gender

```{r}
plt("gender", "gender")
```

### Education Level
```{r}
plt("education_level", "education")
```

### Age

```{r}
plt("age", "age")
```

### Snore Frequency

```{r}
plt("snore_freq", "snore frequency")
```

### Feel sleepy frequency

```{r}
plt("feel_sleepy_freq", "feel sleepy frequency")
```

### Whether the sleep duration is sufficient

```{r}
plt("if_enough_sleep", "Whether the sleep duration is sufficient")
```

### Depression screening score

```{r}
plt("depression_score", "depression score")
```




### Drink frequency last year

```{r}
plt("drink_freq_last_year", "drink frequency last year")
```

### 4-5 Drinks frequency

```{r}
plt("more_drink_freq", "4-5 Drinks frequency")
```

### Number of times having 4-5 drinks in the last 30 days

```{r}
plt("more_drink_30days", "4-5 drinks in the last 30 days")
```


## Distribution of whether sufficient sleep respect to important factors {.tabset}


### Age

```{r}
p3 <- ggplot(all_p_mr, aes(x=age, fill = if_enough_sleep)) +
  geom_histogram(position = "fill")

ggplotly(p3) %>% 
  layout(barmode = "overlay",
    hovermode = "compare")
```


### Depression score

```{r}
p4 <- ggplot(all_p_mr, aes(x=depression_score, fill = if_enough_sleep)) +
  geom_histogram(position = "fill") + 
  labs(title = "Proportion of Adequate Sleep by depression score",
       x = "Score",
       y = "Proportion")
dep_by_age_plot <- ggplotly(p4) %>% 
  layout(barmode = "overlay",
    hovermode = "compare")
ggplotly(dep_by_age_plot)
```

Note that a small number of participants are having a high depression screening score, which explains the chaotic ratio of the distribution whether getting enough sleep.

### Gender

```{r}
p5 <- ggplot(all_p_mr, aes(x=gender, fill = if_enough_sleep)) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Adequate Sleep by Gender",
       x = "Gender",
       y = "Proportion") +
  theme_minimal() +  # Example theme, change as needed
  theme(legend.position = "bottom")  # Adjust legend position
ggplotly(p5) %>% 
  layout(barmode = "overlay",
    hovermode = "compare")
```


### How often feeling sleepy

```{r}
p6 <- ggplot(all_p_mr, aes(x=feel_sleepy_freq, fill = if_enough_sleep)) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Adequate Sleep by Sleep feeling",
       x = "Feeling sleepy frequency",
       y = "Proportion") +
  coord_flip()+
  theme(legend.position = "bottom")  # Adjust legend position
ggplotly(p6) %>% 
  layout(barmode = "overlay",
    hovermode = "compare")
```


### whether having trouble sleeping

```{r}
p7 <- ggplot(all_p_mr, aes(x=if_trouble_sleeping, fill = if_enough_sleep)) +
  geom_bar(position = "fill") + 
  labs(title = "Proportion of Adequate Sleep by whether having trouble sleeping",
       x = "If have trouble sleeping",
       y = "Proportion")
plotly_p7 <- ggplotly(p7) %>% 
  layout(barmode = "overlay",
    hovermode = "compare")
ggplotly(plotly_p7)
```




## Table {.tabset}

```{r}
skim(all_p_mr)
```


