---
title: "Model statistics"
author: "Xinxiang Gao"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---
```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(widgetframe)
library(tidytext)

# INITALIZE CODE CHUNK OPTIONS
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{r}
source("model.R")
```

## GLM (logistic regression) with AIC

| Variable | Estimate | Pr(>\|z\|) |
|---------------------------------------------------------|----------|--------------|
| gender Male | -0.623 | < 0.001* |
| education_level College graduate or above | 0.257 | 0.091 |
| education_level High school graduate/GED or equivalent | 0.003 | 0.983 |
| education_level Less than 9th grade | 0.625 | 0.015*|
| education_level Some college or AA degree | -0.053 | 0.710 |
| age | -0.014 | < 0.001*|
| snore_freq Never | 0.276 | 0.015*|
| snore_freq Occasionally - 3-4 nights a week | 0.084 | 0.462 |
| snore_freq Rarely - 1-2 nights a week | 0.184 | 0.078 |
| feel_sleepy_freq Never | 0.876 | < 0.001* |
| feel_sleepy_freq Often- 5-15 times a month | 0.423 | 0.005* |
| feel_sleepy_freq Rarely - 1 time a month | 0.902 | < 0.001* |
| feel_sleepy_freq Sometimes - 2-4 times a month | 0.775 | < 0.001* |
| depression_score | -0.019 | 0.038* |

* indicates the p-value suggests the statistical significance of the coefficient

AIC: 3953.6

## Random forest

```{r}
varImpPlot(
  rf,
  cex.lab = 1.5, cex.axis = 2, cex = 1.3,
  n.var = 13, col = "red4", main = "Relative importance in random forest", pch = 16)
```

## XGBoost (Extreme gradient boost)

```{r}
plot(xgb_model)
varimp <- varImp(xgb_model)
ggplot(varimp) +
  ggtitle("Relative importance in XGBoost")
```

